Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-15 document/ms.tex
fixing stupid type in transit probability... the code is right though

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, it has been shown that these populations permit an
uncertain extrapolated estimate of the probability or rate at which Sun-like
stars host true Earth analogs.
We relax these assumptions by modeling the occurence rate as a
multidimensional Gaussian process and find that the rate density of Earth
analogs is about two percent with large uncertainty.

@@

The detection efficiency also depends on the geometric transit probability
$R_\star/a$.
Since we're modeling the distribution in the period--radius plane, we need to
compute the transit probability marginalized over stellar radius and mass.
This marginalized distribution scales only with the period of the orbit as
$\propto \period^{-2/3}$.
In theory, this marginalization should be over the true distribution of these
parameters in the selected stellar catalog but we'll approximate it by
the empirical distribution; a reasonable approximation given the size of the
dataset.
At a period of 10 days, the median transit probability in the selected sample
of stars is $5\%$ so we model the transit probability\footnote{We're using the
letter $Q$ to indicate probabilities since we're already using $P$ to mean
period.} as a function of period as
$\transitprob (\period) = 0.05\,(\period-10)^{-2/3}$.

>>>>

An impressive number of exoplanets have been discovered and it is of great
scientific interest to study what we can say about the exoplanet occurence
rate function given these discoveries.
Every catalog is censored by non-trivial selection effects and detection
efficiencies, and the measurements of the system properties (period, radius,
\etc) are noisy.
We present a general hierarchical probabilistic framework for making justified
inferences about the occurrence rate of exoplanets taking into account survey
completeness \emph{and observational uncertainties}.
By applying our method to synthetic catalogs, we demonstrate that it produces
more accurate estimates of the underlying occurence rate than the standard
inverse-detection-efficiency weighting procedure.
When applied to an existing catalog of small planet candidates around G dwarfs
(Petigura \etal\ 2013), our method confirms previous measurements of a
flattening radius distribution near Earth's radius.
Although no true extra-solar Earth analog is known, hundreds of planets have
been found around Sun-like stars that are either Earth-sized but on shorter
periods, or else on year-long orbits but somewhat larger.
Under strong assumptions, it has been shown that these populations permit an
uncertain extrapolated estimate of the probability or rate at which Sun-like
stars host true Earth analogs.
We relax these assumptions by modeling the occurence rate as a
multidimensional Gaussian process and find that the rate density of Earth
analogs is about two percent (per natural logarithmic bin in period and
radius) with large uncertainty.

@@

The detection efficiency also depends on the geometric transit probability
$R_\star/a$.
Since we're modeling the distribution in the period--radius plane, we need to
compute the transit probability marginalized over stellar radius and mass.
This marginalized distribution scales only with the period of the orbit as
$\propto \period^{-2/3}$.
In theory, this marginalization should be over the true distribution of these
parameters in the selected stellar catalog but we'll approximate it by
the empirical distribution; a reasonable approximation given the size of the
dataset.
At a period of 10 days, the median transit probability in the selected sample
of stars is $5\%$ so we model the transit probability\footnote{We're using the
letter $Q$ to indicate probabilities since we're already using $P$ to mean
period.} as a function of period as
\begin{eqnarray}
\transitprob (\period) &=&
    0.05\,\left[\frac{\period}{10\,\mathrm{days}}\right]^{-2/3} \quad.
\end{eqnarray}
